---
title: "SA1 TIME SERIES Q14"
author: "Frances Aneth Rosales"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true        
  html_document:
    code_folding: hide
    toc: true        
---

```{r setup , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readxl)
library(forecast)
library(tseries)
library(tidyr)
library(dplyr)
library(afex)
library(ggplot2)
library(car)
library(ggplot2)
library(zoo)
library(lubridate)
library(flextable)
library(readr)
library(car)
library(emmeans)

```

<class style="font-size: 15px;">
Import </class>
```{r 1st, include=FALSE , warning=FALSE}
ratexplo_data <- read_csv("C:/Users/asus/Documents/ALL FEU FILES/FEU FOLDER 7/APPLIED MULTIVARIATE/SA1/RatExploration_csvFile.csv") 

ratexplo_data <- ratexplo_data[, -4]

ratexplo_data <- ratexplo_data %>%
  filter(!is.na(ID) & !is.na(Stimuli) & !is.na(Time)) %>%
  mutate(Time = as.numeric(Time),
         Stimuli = as.factor(Stimuli))

ratexplo_data$Stimuli <- as.factor(ratexplo_data$Stimuli)

```

\newpage
\section{READING THE DATA}

<class style="font-size: 15px;">
Import </class>

```{r show1st  }

print(head(ratexplo_data))

```




\newpage
\section{1. Summary of ANOVA: Effect of Oil Type on Mileage}
```{r Normality, include=FALSE}
anova_results <- aov(Time ~ Stimuli, data = ratexplo_data)

anova_summary <- summary(anova_results)

```

```{r Normalityshow  }
print(anova_summary)
```
\subsubsection{Findings:} 

According to our ANOVA results, we saw that the time spent in three different treatments is \textbf{statistically significant}. As such, with an \textbf{F-value of 62.09} and a highly significant \textbf{p-value of 6.53e-12}, this shows that there's a probability that one of the different treatments such as \textbf{shapes, patterns, or pictures} implicates that rats exhibit various levels of exploratory behavior.

```{r Independence1show  }

if (anova_summary[[1]]["Pr(>F)"][1, 1] < 0.05) {
  cat("There is a significant difference in time spent among the treatment conditions.\n")
} else {
  cat("There is no significant difference in time spent among the treatment conditions.\n")
}
```
\newpage

\section{2. POST-HOC}
```{r Homogeneityshow  }
if (anova_summary[[1]]["Pr(>F)"][1, 1] < 0.05) {
  posthoc_results <- TukeyHSD(anova_results)
  print(posthoc_results)
}
```

\subsubsection{Findings:}
According to our \textbf{Tukey post-hoc analysis}, there are also \textbf{significant differences} in the time spent on three different treatments. The results we obtained show an \textbf{average increase of 1.07 seconds} in the \textit{Picture} condition compared to the \textit{Pattern} condition, with a \textbf{p-value of 0.00034}. On the other hand, \textit{Shape} decreased by \textbf{1.64 seconds} compared to \textit{Pattern} with a \textbf{p-value of 0.0000004}. Lastly, there is a decrease of \textbf{2.70 seconds} in \textit{Shape} compared to \textit{Picture}, with a \textbf{p-value of 0.0000000}. This implies that the type of visual stimulus indeed \textbf{significantly influences the exploratory behavior} of the rats, with \textit{Picture} being the \textbf{most effective} and \textit{Shape} appearing to be the \textbf{least effective}.



\newpage


\section{3. Boxplot Time vs Treatment Condition }

```{r 2waysummaryshow  }

ggplot(ratexplo_data, aes(x = Stimuli, y = Time)) +
  geom_boxplot() +
  labs(title = "Exploration Time by Treatment Condition",
       x = "Treatment Condition",
       y = "Seconds Spent Exploring") +
  theme_minimal()


```
\subsubsection{Findings:}
Our boxplot shows that the \textbf{"Picture"} treatment is associated with the \textbf{most time spent} compared to the other two treatments, with a median of approximately \textbf{4 seconds}. However, \textbf{"Shape"} shows the \textbf{lowest exploration} among the three treatments, with a median of less than \textbf{2 seconds}. \textbf{"Pattern"}, on the other hand, falls in between the two treatment conditions.   


\newpage

\section{ANALYSIS AND CONCLUSION}

In this scenario, we are required to analyze the data using one of the \textbf{ANOVA types} to compare the number of seconds the rats spent exploring the experimental chamber with the images. Additionally, we need to determine if there is a \textbf{significant difference} in time spent between the three treatment conditions: \textit{shapes}, \textit{patterns}, and \textit{pictures}.

Among the results, our analysis shows that the three treatments of visual stimulus \textbf{indeed} contributed a significant impact on their exploration behavior. According to our findings, \textbf{Pictures} show the \textbf{most exploratory behavior} in rats, followed by \textit{patterns} and lastly, \textit{shapes}. This points to the possibility that various appearances, some more engaging than others, can affect rats' exploratory behavior.
